---
title: "Getting acquainted with Mastodon -- Accounts"
author: "Karsten W."
date: "2022-10-30"
draft: TRUE
output: html_document
vignette: >
  %\VignetteIndexEntry{Getting to know YOU on Mastodon}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

First, create an account on an Mastodon server of your choice. Then, provide
your login credentials to R. You might consider putting it into your .Rprofile. 

```{r}
#options(
#	msocial.instance="https://berlin.social/",
#	msocial.email="me@email.com",
#	msocial.passw="secret"
#)
```

Next, load the need libraries.

```{r}
library(msocial)
library(mastodon) # devtools::install_github('ThomasChln/mastodon')
token <- login(
	getOption("msocial.instance"),
	getOption("msocial.email"), 
	getOption("msocial.passw")
)
```

Now we are ready to use the API and find some posts that use the #rstats tag. We are interested in their instances only. We can fetch some information on the instances:

```{r}
dat <- get_hashtag(token, 'rstats', local = FALSE, n=10000)
users <- c(users, sapply(dat$account, function(x) x$acct))
instances <- unique(gsub("^.*@", "https://", users))
stats <- do.call(rbind, lapply(instances, get_instance_stats))
head(stats)
```

Also, we can fetch public user profile data from each instance:

```{r}
get_instance_users(instances[1])
```

